 Casting multiple shadows: high-dimensional interactive data visualisation with tours and embeddings

### 2020-06-23 paper.Rmd
rejig
tidy up case study no. 2 section

While there has been much work on the algorithmic details of the aforementioned
embedding methods, there has been relatively few tools designed to assist users
to interact with these techniques and assist them in making comparisons between
embeddings and performing the aforementioned cluster orientation tasks. Several
interactive interfaces have been proposed for evaluating or using embedding
techniques: the Sleepwalk interface provides a click and highlight
visualisation for colouring points in an embedding according to their distance
in the original high-dimensional space [@Ovchinnikova2019-gf]. The work by
@Pezzotti2017-cz provides a user guided and modified form of the t-SNE
algorithm, that allows users to modified optimisation parameters in real-time.
Similarly, the embedding projector is a web interface to running UMAP, t-SNE or
PCA live in the browser and provides interactions to colour points, and
highlights nearest neighbours [@Smilkov2016-hp].

@@

A common analysis task in single cell studies is performing clustering to
identify groupings  of cells with similar expression profiles. Analysts
in this area generally use non linear DR methods for verification and identifaction
of clusters and developmental trajectories (i.e., case study 1). For clustering,
workflows the primary task is verify the existence of clusters and then begin to 
identify the clusters as cell types using the expression of "known" 
marker genes. The task becomes harder if the goal is discovery of new cell
types

Here we describe a liminal view of mouse single cell data from @Macosko2015-ot
following the workflow described in @Amezquita2020-at. To begin we obtained
the count matrix from the `scRNAseq

>>>>

While there has been much work on the algorithmic details of the aforementioned
embedding methods, there has been relatively few tools designed to assist users
to interact with these techniques and assist them in making comparisons between
embeddings and performing the aforementioned cluster orientation tasks. Several
interactive interfaces have been proposed for evaluating or using embedding
techniques: the Sleepwalk interface provides a click and highlight
visualisation for colouring points in an embedding according to their distance
in the original high-dimensional space [@Ovchinnikova2020-sy]. The work by
@Pezzotti2017-cz provides a user guided and modified form of the t-SNE
algorithm, that allows users to modified optimisation parameters in real-time.
Similarly, the embedding projector is a web interface to running UMAP, t-SNE or
PCA live in the browser and provides interactions to colour points, and
highlights nearest neighbours [@Smilkov2016-hp].

@@

A common analysis task in single cell studies is performing clustering to
identify groupings  of cells with similar expression profiles. Analysts
in this area generally use non linear DR methods for verification and identifaction
of clusters and developmental trajectories (i.e., case study 1). For clustering,
workflows the primary task is verify the existence of clusters and then begin to 
identify the clusters as cell types using the expression of "known" 
marker genes. Here a 'faithful' a embedding should ideally preserve the topology
of the data; cells that correspond to a cell type should lie in the same
neighbourhood in high-dimensional space. In this case study we use our 
spatial and linked brushing approaches to look at neighbourhood preservation
(which can be seen via the tour) and to look at the affect of applying different
feature selection techniques and distance metrics.

First, we downloaded the raw mouse retinal single cell RNA-seq data from
@Macosko2015-ot using the `scRNAseq` Bioconductor package [@scRNAseq-d]. We
have followed a standard workflow for pre-processing and normalizing this data 
(described by @Amezquita2020-at): we performed QC using the the `scater`
by removing cells with high proportion of mitocondrial gene expression and low
numbers of genes detected, we log-transformed and normalised the expression 
values and  finally selected highly variable genes (HVGs) using `scran` 
[@McCarthy2017; @Lun2016]. The top ten percent of HVGs were used to subset
the normalised expression matrix and compute PCA using the first 25 components.
Using the PCs we built a shared nearest neighours graph (with k = 10) and 
used Louvain clustering to generate clusters. To check and verify the 
clustering we construct a liminal view. Because most off the variablity in
expression is explained by the first 5 PCs, we tour those. The t-SNE view
is computed from all PCs

- compare to just using the HVGs them selves 
- neighbourhood based brushing to check how diffuse or distorted clusters are
- different distance metrics?
